Vagueness, Noise, and Signaling

### 2015-08-12 paper/version_04/paper.tex
explain new dynamics

Signaling agents can adapt their dispositions to act, given some feedback about their past
success, in multiple ways. Usually, we would assume that changes in behavior should, at least
on average, lean towards increasing chances of communicative success. Such a population
of signaling agents and their behavioral adaptations can be described at different levels of
abstraction. At the level of individual agents, we can describe a more or less idealized
process of how each agent adapts dispositions for future actions based on detailed descriptions
of information available to the agent, e.g., about the game, the other players' behavior and
the like. More abstractly, at the level of a population of agents, we can describe how
average behavioral dispositions will evolve. The latter population-level perspective abstracts
over petty stochastic fluctuations and zooms in on the general tendency or direction of
evolution that ensues from some (often: simplified) behavior at the agent-level.

We are interested here in exploring how a tendency to favor communicatively successful
strategies is affected by imprecise perception and realization of states. We therefore look at
a population-level dynamic that describes the most likely evolutionary path of a population of
signaling agents who imitate other agents' behavior (proportional to how successful it is) but
may occasionally confuse states for one another (proportional to how similar they are). We call
our dynamic \emph{imprecise conditional imitation dynamic} (\icid), because it derives from an
update protocol based on conditional imitation and allows for imprecise perception and
realization of states. In the special limiting case where imprecision vanishes, the process is
just the well-known replicator dynamic. To appreciate the \icid and its motivation, we first
briefly review the replicator dynamic and its derivation from conditional imitation of
successes.

@@

Fix a signaling game with finite states $\States$, messages $\Messgs$ and acts $\Acts$. Let
$\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
$\Util_{\sen,\rec} \mycolon \States \times \Messgs \times \Acts \rightarrow \mathds{R}$ the
sender's and receiver's utility functions. A \emph{behavioral strategy} is a function that maps
an agent's choice points to a probability distribution over available choices.\footnote{Our
  focus is on behavioral strategies, not \emph{mixed strategies}, i.e., probability
  distributions over functions from each choice point to an act, such as
  $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume that agents can
  adapt their behavior locally, so to speak, i.e., independently at each choice point. Our
  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
  numerical simulations. But it also seems the more plausible choice for imitation-based update
  protocols: agents only observe how, on some occasion, some other agent behaved in \emph{one}
  particular situation, not how that agent would behave in \emph{all} relevant choice
  situations; they imitate the use of a single word, so to speak, not a whole lexicon. (See
  \citet{Cressman2003:Evolutionary-Dy} for more on the difference between dynamics on mixed or
  behavioral strategies.)} The sender's behavioral strategies are functions
$\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a
probability of each message $\messg \in \Messgs$ being sent; the receiver's are functions
$\Rstrat \in \Delta(\Acts)^\Messgs$, thus mapping each message $\messg \in \Messgs$ to a
probability of each state $\state \in \States$ being interpreted by the receiver given
$\messg$. Although behavioral strategies are probabilistic, we may conveniently imagine that
every individual agent has a non-probabilistic strategy nonetheless. If the population is large
enough, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the probability
that a randomly sampled sender would have a strategy that responds with $\messg$ to $\state$.

The \emph{expected utility} of choices at each choice point is:
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
    \States} \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
This means that the expected utility for the sender of choosing message $\messg \in \Messgs$,
given tate $\state \in \States$ and a receiver strategy $\Rstrat$ is the sum, for each act
$\act \in \Acts$, of the probability that the receiver will choose act $\act$ given message
$\messg$ times the actual utility of using act $\act$ for state $\state$.  The expected utility
for the receiver of choosing act $\act \in \Acts$, given a message $\messg \in \Messgs$ and a
sender strategy $\Sstrat$ is the sum, for each state $\state \in \States$, of the prior
probability of state $\state$ occurring, times the probability the sender will choose message
$\messg$ given state $\state$, times the actual utility of using act $\act$ for state $\state$.

% The \emph{fitness} of a behavioral strategy at a choice point is the
% frequency-weighted average of expected utilities of each choice, given
% the opponent's strategy:
% \begin{align*}
%   \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
% \EU(\messg, \state,\Rstrat) \\
% \Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
% \cdot \EU(\act, \messg,\Sstrat)\,.
% \end{align*}

The discrete-time replicator dynamic maps current strategies $\Sstrat$ and $\Rstrat$ to future
strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$ in such a way that changes in frequency are
proportional to expected utilities:
\begin{align}
  \label{eq:RDDiscrete}
  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
    \EU(\messg, \state,\Rstrat)} {\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot
   \EU(\messg', \state,\Rstrat)} \\
    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
    \EU(\act, \messg,\Sstrat)} {\sum_{\act'} \Rstrat(\act' \probbar \messg)
    \cdot \EU(\act', \messg,\Sstrat)}  \,. \nonumber
\end{align}
Intuitively put, this definition updates the average frequency with which each choice (message
or act) will occur, independently at each choice point (state or message). For a given choice
point, say a state $\state$, the probability of seeing $\messg$ played by an average agent in
the population \emph{before} the update is $\Sstrat{\messg \probbar \state}$. \emph{After} the
update it will be proportional to the frequency before the update times the expected utility of
$\messg$ at state $\state$. The denominator is then a normalizing constant to make sure that we
obtain probabilities that sum to one from this. In particular, the denominator in the
definition above captures the average expected utility of all choices at the relevant choice
point.

@@

The replicator dynamic is an abstract population-level dynamic that describes the mean,
to-be-expected change in behavioral dispositions in a population of signalers. There are
several ways of deriving the replicator dynamic from agent-level processes of behavioral
adaptation. We focus here on one of the simplest: \emph{conditional imitation of success}
\citep[see][for this and many other ways of deriving the replicator
dynamics]{Sandholm2010:Population-Game}. We choose this agent-level derivation, because it is
more convenient than its alternatives when we want to add imprecision in the following. The
main idea is that agents imitate the behavior of another agent at some choice point with a
probability that is proportional to the expected utility of the latter agent's act for that
choice point. Since the sim-max games that we will be looking at have all expected utilities
conveniently lie between 0 and 1, we may even identify the switching probability with the
expected utility directly.

Let's consider sender strategies, as the receiver case is parallel (as long as there is no
noise or imprecision). Revision by imitation of success proceeds as follows. Call
(misleadingly!) the agent who gets a chance to change behavior ``learner'' and the possibly
to-be-imitated agent ``teacher.'' A random learner is drawn from the population and given a
chance to change behavior at choice point $\state$. The probability that that learner plays
$\messg$ is $\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled
teacher does at $\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar
\state)$. The learner then starts using $\messg'$ instead of $\messg$ with probability
$\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and $\messg$ could be the same.)

@@

From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
to:
\begin{align*}
  \dot{\Sstrat}(\messg \probbar \state) & = \Sstrat(\messg \probbar \state) \cdot \EU(\messg,
  \state, \Rstrat) - \Sstrat(\messg \probbar \state) \cdot  \sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
& = \underbrace{\Sstrat(\messg \probbar \state)}_{\text{frequency of $\messg$ at $\state$}}  ( \underbrace{\EU(\messg,
  \state, \Rstrat)}_{\text{EU of $\messg$ at $\state$}} -   \underbrace{\sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat)}_{\text{average
      EU at choice point $\state$}}  ) 
\end{align*}
This is the continuous-time formulation of the replicator dynamics for behavioral
strategies. The discrete time formulation in Equation (\ref{eq:RDDiscrete}) solves this
continuous-time formulation by assuming that discrete update steps are infinitesimally
small:\footnote{Notice that we drop the denominator here, to obtain the non-payoff corrected
  replicator dynamic.}
\begin{align*}
  \dot{\Sstrat}(\messg \probbar \state) & = \RD(\Sstrat)(\messg \probbar \state) -
  \Sstrat(\messg \probbar \state) \\
  & = \frac{\Sstrat(\messg \probbar \state) \cdot \EU(\messg, \state,\Rstrat)} {\sum_{\messg'}
    \Sstrat(\messg' \probbar \state) \cdot
    \EU(\messg', \state,\Rstrat)} - \Sstrat(\messg \probbar \state) \\
  & = \frac{\Sstrat(\messg \probbar \state) \cdot \EU(\messg, \state,\Rstrat) - \Sstrat(\messg
    \probbar \state) \cdot \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
    \state,\Rstrat)} {\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot
    \EU(\messg', \state,\Rstrat)}  \\
  & \propto \Sstrat(\messg \probbar \state) \cdot \left ( \EU(\messg, \state,\Rstrat) - \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
  \state,\Rstrat) \right )
\end{align*}

\bigskip

 Suppose, she currently plays $c$ at $k$. She
observes what a randomly sampled teacher would do at choice point $k$, say $c'$ and switches
from $c$ to $c'$ with switch probability $\text{SWITCH}(c \rightarrow c' , k)$, which we can
here identify with the expected utility $\EU(c',k)$ of $c'$ at $k$. The mean change induced by
offering all agents imitation chances for $k$ is:
\begin{align*}
  \text{mean change of frequency $c$, at $k$} & = \text{probability of switchers $c'
    \rightarrow c$ at $k$ \mygray{(inflow)}} \\
& - \text{probability of switchers $c \rightarrow c'$ at $k$ \mygray{(outflow)}}
\end{align*}
We should spell out the inflow and outflow here with some care, so as to prepare ourselves for
the subsequent addition of imprecise perception and realization of stimuli. So in general, we
would define:
\begin{align*}
  \text{inflow} & = \sum_{c'} \text{prob.~random learner maps $k$ to $c'$} \\
  & \ \ \ \times \text{prob.~learner observes $c$ by random teacher} \\
  & \ \ \ \times \EU(c', k) \\
  \text{outflow} & = \sum_{c'} \text{prob.~random learner maps $k$ to $c$} \\
  & \ \ \ \times \text{prob.~learner observes random teacher play $c'$ at $k$} \\
  & \ \ \ \times \text{SWITCH}(c \rightarrow c' , k)
\end{align*}

Suffice it to look at the case of a sender.

>>>>

Signaling agents can adapt their dispositions to act, given some feedback about their past
success, in multiple ways. Usually, we would assume that changes in behavior should, at least
on average, lean towards increasing chances of communicative success. Such a population of
signaling agents and their behavioral adaptations can be described at different levels of
abstraction. At the level of individual agents, we can picture a more or less idealized process
of how each agent adapts dispositions for future actions based on various pieces of information
available to the agent, e.g., about the game, the other players' behavior and the like. More
abstractly, at the level of a population of agents, we can describe how average behavioral
dispositions will evolve. The population-level perspective abstracts over petty stochastic
fluctuations and zooms in on the general tendency or direction of evolution that ensues from
some (often: simplified) behavior at the agent-level.

To better understand the interaction of selective pressure successful communication and
stochastic noise, we look at a population-level dynamic that describes the most likely
evolutionary path of a population of signaling agents who imitate other agents' behavior but
may occasionally confuse states for one another. We call our dynamic \emph{imprecise
  conditional imitation dynamic} (\icid), because it derives from an update protocol based on
conditional imitation and allows for imprecise perception and realization of states. In the
special limiting case where imprecision vanishes, the process is just the well-known replicator
dynamic \citep{TaylorJonker1978:Evolutionary-St}. To appreciate the \icid and its motivation,
we first briefly review the replicator dynamic and its derivation from conditional imitation of
successes.

@@

Fix a signaling game with finite states $\States$, messages $\Messgs$ and acts $\Acts$. Let
$\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
$\Util_{\sen,\rec} \mycolon \States \times \Messgs \times \Acts \rightarrow \mathds{R}$ the
sender's and receiver's utility functions. A \emph{behavioral strategy} is a function that maps
an agent's choice points to a probability distribution over available choices.\footnote{Our
  focus is on behavioral strategies, not \emph{mixed strategies}, i.e., probability
  distributions over functions from each choice point to an act, such as
  $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume that agents can
  adapt their behavior locally, so to speak, i.e., independently at each choice point. Our
  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
  numerical simulations. But it also seems the more plausible choice for imitation-based update
  protocols: agents only observe how, on some occasion, some other agent behaved in \emph{one}
  particular situation, not how that agent would behave in \emph{all} relevant choice
  situations; they imitate the use of a single word, so to speak, not a whole lexicon. (See
  \citet{Cressman2003:Evolutionary-Dy} for more on the difference between dynamics on mixed or
  behavioral strategies.)} The sender's behavioral strategies are functions
$\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a
probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's are
functions $\Rstrat \in \Delta(\Acts)^\Messgs$, thus mapping each message $\messg \in \Messgs$
to a probability of each act $\act \in \Acts$ being chosen in response to $\messg$. Although
behavioral strategies are probabilistic, we may conveniently imagine that every individual
agent has a non-probabilistic strategy nonetheless. If the population is large enough, the
number $\Sstrat(\messg \probbar \state)$, for instance, is then the probability that a randomly
sampled sender would have a strategy that responds with $\messg$ to $\state$.

The \emph{expected utility} of choices at each choice point is:
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
    \States} \frac{\Pr(\state) \cdot \Sstrat(\messg \probbar \state)}{\sum_{\state'} \Pr(\state') \cdot \Sstrat(\messg \probbar \state')} \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
outcomes, weighted by how likely they are to occur. For instance, the expected utility for the
sender of choosing message $\messg \in \Messgs$, given tate $\state \in \States$ and a receiver
strategy $\Rstrat$ is the sum, for each act $\act \in \Acts$, of the probability that the
receiver will choose act $\act$ given message $\messg$ times the actual utility of using act
$\act$ for state $\state$, given $\state$ and $\messg$.

The continuous-time replicator dynamic for behavioral strategies describes the mean rate of
change in the proportion of each choice, independently for each choice point:
\begin{align}
  \label{eq:RDContinuous}
  \dot{\Sstrat}(\messg \probbar \state) & = \underbrace{\Sstrat(\messg \probbar
    \state)}_{\text{frequency of $\messg$ at $\state$}}  \Big ( \underbrace{\EU(\messg,
  \state, \Rstrat)}_{\text{EU of $\messg$ at $\state$}} -   \underbrace{\sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat)}_{\text{average
      EU at choice point $\state$}} \Big ) \\
  \dot{\Rstrat}(\act \probbar \messg) & = \Rstrat(\act \probbar \messg) \thinspace \Big ( \EU(\act,
  \messg, \Sstrat) -  \sum_{\act'} \Rstrat(\act' \probbar \messg) \cdot \EU(\act',
  \messg,\Sstrat)  \Big ) \nonumber
\end{align}
The rate of change in proportion of $\messg$ choices at choice point $\state$, for instance, is
given by the difference of $\messg$'s expected utility at $\state$ and the average expected
utility at $\state$, times the probability of seeing $\messg$ at $\state$ in the first place.

For numerical simulations, it is often more convenient to work with a discrete-time
formulation, which maps current strategies $\Sstrat$ and $\Rstrat$ to strategies $\RD(\Sstrat)$
and $\RD(\Rstrat)$ that describe the population at the next discrete time step. Like the
continuous-time version, the discrete-time replicator dynamics tracks changes in frequency of
choices as proportional to their expected utilities
\citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
  utilities are always positive.}
\begin{align}
  \label{eq:RDDiscrete}
  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
    \EU(\messg, \state,\Rstrat)} {\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot
   \EU(\messg', \state,\Rstrat)} \\
    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
    \EU(\act, \messg,\Sstrat)} {\sum_{\act'} \Rstrat(\act' \probbar \messg)
    \cdot \EU(\act', \messg,\Sstrat)}  \,. \nonumber
\end{align}
Intuitively put, this definition updates the average frequency with which each choice (message
or act) will occur, independently at each choice point (state or message). For a given choice
point, say a state $\state$, the probability of seeing $\messg$ played by an average agent in
the population \emph{before} the update is $\Sstrat(\messg \probbar \state)$. \emph{After} the
update it will be proportional to the frequency before the update times the expected utility of
$\messg$ at state $\state$. The denominator is then a normalizing constant to make sure that we
obtain probabilities that sum to one. In particular, the denominator in the
definition above captures the average expected utility of all choices at the relevant choice
point.

@@

The replicator dynamic is an abstract population-level dynamic that describes the mean expected
change of behavioral dispositions in a population of signalers. There are several ways of
deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
here on one of the simplest: \emph{conditional imitation of success} \citep[see][for this and
many other derivations]{Sandholm2010:Population-Game}. We choose this agent-level derivation,
because it is more convenient than its alternatives when we want to add imprecision in the
following. The main idea is that agents imitate the behavior of another agent at some choice
point with a probability that is proportional to the expected utility of the latter agent's act
for that choice point. Since the sim-max games that we will be looking at in
Section~\ref{sec:exploring-rdd} have positive utilities, we may even identify the switching
probability with the expected utility directly.

Let's consider sender strategies, as the receiver case is parallel (as long as there is no
noise or imprecision). Revision by imitation of success proceeds as follows. Call
(misleadingly!) the agent who gets a chance to change behavior ``learner'' and the possibly
to-be-imitated agent ``teacher.'' A random learner is drawn from the population and given a
chance to change behavior at choice point $\state$. The probability that our learner plays
$\messg$ is $\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled
teacher does at $\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar
\state)$. The learner then starts using $\messg'$ instead of $\messg$ with probability
$\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and $\messg$ could be the same.)

@@

From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
to:
\begin{align*}
  \dot{\Sstrat}(\messg \probbar \state) & = \Sstrat(\messg \probbar \state) \cdot \EU(\messg,
  \state, \Rstrat) - \Sstrat(\messg \probbar \state) \cdot  \sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
& = \Sstrat(\messg \probbar \state) \cdot \Big ( \EU(\messg, \state,\Rstrat) - \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
  \state,\Rstrat) \Big )\,,
\end{align*}
which is the continuous-time formulation of the replicator dynamic in (\ref{eq:RDContinuous}).

\subsection{Imitation of success with noise}
\label{sec:imit-succ-with}

The above derivation of the replicator dynamic assumes that agents can discriminate choices and
choice points perfectly. Let's dispense with that assumption. With an eye towards sim-max
games, we will assume that receivers choose interpretation actions $\Acts = \States$ and that
any two states may have a chance of being confused for one another, proportional to how similar
they are. This will affect the way that agents realize the (pure) strategies that they carry
and also how learners observe teacher choices and the probability with which they switch to the
observed teacher choice.

Agents carry pure dispositions to act, but noise can affect their realization. As a sender,
every agent maps states to messages: these are subjectively perceived states, and no longer
necessarily also the actually occurring state. As a receiver, every agent maps messages to
state interpretations: these are intended interpretations that need not always be faithfully
realized. This means that behavioral strategies $\Sstrat \in (\Delta(\Messgs))^\States$ and
$\Rstrat \in (\Delta(\States))^\Messgs$ represent the average proportions of \emph{actual}
behavioral dispositions in the population, the realization and observation of which can be
distorted by confusion of similar states.

If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
probability that a given agent observes state $\mystate{o}$. States $\mystate{a}$ and
$\mystate{o}$ may be identical. In the limiting case, $P_o(\mystate{o} \probbar \mystate{a})$
could be one only if $\mystate{a}$ and $\mystate{o}$ are identical. For the more general case,
we should treat $P_o(\mystate{o} \probbar \mystate{a})$ as a primitive of the model. Similarly,
if a given receiver intends to select interpretation $\mystate{i}$, we will assume that the
probability with which state $\mystate{r}$ is realized is
$P_r(\mystate{r} \probbar \mystate{i})$. Section~\ref{sec:exploring-rdd} gives a concrete
specification of $P_o$ and $P_r$ in terms of similarity of states.

A single round of play is then governed by five pieces of stochastic information, where
previously there were only three. At the outset, the actual state is chosen with probability
$P_a(\mystate{a})$. A randomly sampled sender observes $\mystate{o}$ with probability
$P_o(\mystate{o} \probbar \mystate{a})$ and so we expect $\messg$ with probability
$\Sstrat(\messg \probbar \mystate{o})$. (Senders respond to observations, not the actual
state!) As perception of messages is flawless, we expect a randomly sampled receiver to intend
to realize interpretation $\mystate{i}$ with probability
$\Rstrat(\mystate{i} \probbar \messg)$.  The actually realized interpretation will be
$\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$.

Expected utilities of choices at choice points should take into account as well that actual
states need not be observed states, and intended interpretations need not be realized
interpretations. First, note that 
\begin{align*}
  P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) = \frac{P_a(\mystate{a}) P_o(\mystate{o}
    \probbar \mystate{a})}{\sum_{\state}P_a(\state) P_o(\mystate{o}
    \probbar \state)}
\end{align*}
is the probability that $\mystate{a}$ is actual if $\mystate{o}$ is observed. The probability
that a random sender produces $\messg$ when the actual state is $\mystate{a}$ is:
\begin{align*}
  P_{\Sstrat}(\messg \probbar \mystate{a}) = \sum_{\mystate{o}} P_o(\mystate{o} \probbar
  \mystate{a}) \Sstrat(\messg \probbar \mystate{o})\,. 
\end{align*}
The probability that the actual state is $\mystate{a}$ if a random sender produced $\messg$ is:
\begin{align*}
  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) = \frac{P_a(\mystate{a})
  P_{\Sstrat}(\messg \probbar \mystate{a})}{\sum_{\state} P_a(\state)
  P_{\Sstrat}(\messg \probbar \state)}\,. 
\end{align*}
The probability that $\mystate{r}$ is realized by a random receiver in response to
message $\messg$ is:
\begin{align*}
  P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
  \mystate{i}) \Rstrat(\mystate{i} \probbar \messg)\,.
\end{align*}
This lets us capture the expected utilities for observed states (sender) and intended
interpretations (receiver) by taking into consideration what the likely actual states and
realized interpretations will be:
\begin{align*}
  \EU(\messg , \mystate{o}, \Rstrat) & = \sum_{\mystate{a}}  P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \sum_{\mystate{r}}  P_\Rstrat(\mystate{r} \probbar
  \messg) \utils(\mystate{a}, \messg, \mystate{r})\,, \\
  \EU(\mystate{i} , \messg, \Sstrat) & = \sum_{\mystate{a}}
  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \sum_{\mystate{r}}
  P_{r}(\mystate{r} \probbar \mystate{i})  \utils(\mystate{a}, \messg, \mystate{r})\,.
\end{align*}
If the conditional probabilities $P_o$ and $P_r$ are trivial, i.e., assign probability 0 to the
observability and realizability of non-identical states, above definitions reduce to the
previous definitions of expected utilities. This also legitimates the overload of notation.

Presence of potential imprecision in the form of non-trivial $P_o$ and $P_r$ will also affect
the dynamic that ensues from imitation of successes. Since imprecision works differently on
senders and receivers (the former confuse choice points, the latter confuse choices), we need
to look separately at each case.

As before, suppose that senders receive a chance to change their behavior independently for a
given choice point. In the present case, this would be a chance to change how to respond to a
subjectively perceived state $\mystate{o}$, which need not be the actual one. We must therefore
take into account what the probability is that, given that the learner observed $\mystate{o}$,
he will observe a randomly sampled teacher play some $\messg$ or other. The probability
$P_o(\messg \probbar \mystate{o})$ that some teacher plays $\messg$ in the same situation in
which some learner observes $\mystate{o}$ is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as
defined above):
\begin{align*}
  P_o(\messg \probbar \mystate{o}) = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \myts P_{\Sstrat}(\messg \probbar \mystate{a})\,.
\end{align*}
The probability $P(\messg' \rightarrow \messg, \state)$ that a randomly sampled learner switches
from $\messg'$ to $\messg$ in subjectively perceived state $\mystate{o}$ is therefore:
\begin{align*}
  P(\messg' \rightarrow \messg, \mystate{o}) = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
    \mystate{o})}_{\text{learner plays $\messg'$ at $\mystate{o}$}} \cdot
  \underbrace{P_{o}(\messg \probbar \mystate{o})}_{\text{observe teacher play $\messg$}} \cdot
  \underbrace{\EU(\messg, \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
\end{align*}
The mean change to the proportion of $\messg$ choices at state $\state$ are then as before:
\begin{align*}
  \dot{\Sstrat}(\messg \probbar \mystate{o}) & = P(\messg' \rightarrow m, \mystate{o}) - P(\messg
  \rightarrow m', \mystate{o}) \\
  & = \sum_{\messg'} \Sstrat(\messg' \probbar \mystate{o}) \myts P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \mystate{o}) \myts
  P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat) \\
  & = P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \Sstrat(\messg \probbar \mystate{o}) \myts \sum_{\messg'} P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat)
\end{align*}
This has an elegant and practical discrete-time solution:\footnote{To see how the discrete-time formulation
  gives rise to the continuous-time formulation, let's assume that update steps are
  infinitesimally small, so that
  \begin{align*}
    \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
    \probbar \state)\\ 
    &= \frac{P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat)
      - \Sstrat(\messg \probbar \state) \sum_{\messg'} P_o(\messg' \probbar \state) \myts
      \EU(\messg', \state, \Rstrat) }{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts
      \EU(\messg', \state, \Rstrat)}\,.
  \end{align*}
By dropping the nominator, which is constant for all $\messg$ for fixed
$\state$, we obtain the above continuous-time formulation.}
\begin{align*}
  \Sstrat'(\messg \probbar \state) = \frac{P_o(\messg \probbar \state) \myts \EU(\messg,
  \state, \Rstrat)}{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts \EU(\messg',
  \state, \Rstrat)}\,.
\end{align*}

The case of the receiver is mostly analogous. Presented with an update opportunity for choice
point $\messg$, a learner will observe a random teacher choose interpretation $\mystate{o}$
with probability:\footnote{We assume that the noise $P_o(\mystate{o} \probbar \mystate{a})$
  that applies to the sender's observations also applies to the receiver's observations when
  trying to imitate other agents' strategies. We could easily make the model more complex and
  introduce another measure of noise for state confusability during receiver attempts to
  imitate. We refrain from it here, because we see no immediate theoretical gain.}
\begin{align*}
  P_o(\mystate{o} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
  \myts P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
  \myts \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
  \mystate{i}) \myts \Rstrat(\mystate{i} \probbar \messg) \,.
\end{align*}
The probability $P(\state' \rightarrow \state, \messg)$ that a randomly sampled learner switches
from using $\state'$ to $\state$ in response to $\messg$ is therefore:
\begin{align*}
  P(\state' \rightarrow \state, \messg) = \sum_{\state'} \underbrace{\Rstrat(\state' \probbar
    \messg)}_{\text{learner plays $\state'$}} \cdot
  \underbrace{P_{o}(\state \probbar \messg)}_{\text{observe teacher play $\state$}} \cdot
  \underbrace{\EU(\state, \messg, \Rstrat)}_{\text{EU teacher choice}}
\end{align*}
This gives rise to continuous-time and discrete-time formulations of resulting dynamics,
parallel to those for the sender case:
\begin{align*}
  \dot{\Rstrat}(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg) \myts
  P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'} \Rstrat(\state \probbar \messg) \myts
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \\
  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar \messg) \myts \sum_{\state'} 
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat)\\
  \Rstrat'(\state \probbar \messg) & = \frac{P_o(\state \probbar \messg) \myts \EU(\state,
  \messg, \Sstrat)}{\sum_{\state'} P_o(\state' \probbar \messg) \myts \EU(\state',
  \messg, \Sstrat)}
\end{align*}

\newpage
